FraCaS: Temporal Analysis

### 2020-02-06 paper2020/paper.tex
expand

Our basic approach is to extend predicates with two time parameters,
one corresponding with the starting time of the action and one with
its stopping time.

walk(john,t0,t1)

\paragraph{time context}

The compositional semantics carry a current time reference (itself an
interval). This current interval propagates through the compositional
interpraetation. By default every interpretation passes the temporal
context to its components --- but some may act on it.

\paragraph{tenses and temporal adverbs}

Tenses are represented syntactically (in GF!) as an attribute of the
clause. In our representation we deal only with present and past
tenses --- we find that FraCas does not exercise (if a variant of the
past is used, the additional information is also carried by adverbs,
in a more specific way.)

If the temporal context was modified by an adverb (or an averbial
clause, such as ``before James swam''), then the tense is ignored.
%
Otherwise, if the tense is the present, then we use a simple
$(now,now)$ interval, with $now$ being a constant.
%
If the tense is the past, then we locally quantify over a time
interval $(t_0,t_1)$, such that $t_1 < now$.

@@

A common theme in the testsuite is reference to previous occurences of an event

P1	Smith left.
P2	Jones left.
P3	Smith left before Jones left.
Q 	Did Jones leave after Smith left?

We have to arrange that the two occurences of ``Jones left'', in P1
and P3 refer to the same time intervals.

@@

\paragraph{(un)repeatable achievements}
Using unicity of action, we can refute 279:

279:
P1	Smith wrote a novel in 1991.
Q 	Did Smith write it in 1992?
H 	Smith wrote it in 1992.

∃x. novel(x)
∀t<NOW.  ∧ write(smith,x,t,t) [the scope for the existential is extended according to our previous analysis of anaphora]

@@

\paragraph{interval adverbs, point adverbs, universal adverbs}
Time adverbs, such as "Since", "always", "never", "in two years", "by
11 am", etc. introduce *local* quantification over a timespan.

time interval:
⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)

Universal:
⟦``always s''⟧(...) = ∀t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)

relative time:
⟦``in two years, s''⟧(t0,tx) = ∃t2.  ⟦s⟧(t0+2years,t2)

@@

\section{Results and evaluation}
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}[hbt]
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#FraCaS
                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
Quantifiers  & 75         & .91        & .96         & .96    & .77  & .53  & .93  \\
             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
Plurals      & 33         & .64        & .82         & .76    & .67  & .52  & .73 \\
             &            &            &             &     &   &   & \ncases{24} \\
Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
             &            &  \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
Comparatives & 31         & .64        & .87         & .56    & .48  & .45  &  -       \\
Temporal     & 75         &  -         &  -          &   -    &   -  &  -   &  -       \\
Verbs        & 8          & .71        & .75         &   -    & -    & -    &  -       \\
Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
             &            &            &             &        &      &      & \ncases {9}  \\ \hline
Total        & 337        & .78        & .89         & .83    & .69  & .50  & .85  \\
             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FraCoq''
    refers to the work of \citet{bernardy:2017}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilizes the first-order automated theorem prover
    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
    to the system presented by \citet{Abzianidze:2015}. A dash
    indicates that no attempt was made for the section. }
  \label{tab:results}
\end{table}
\paragraph{interaction with negation and group readings}
\section{Conclusions and Future Work}
Stergios
\section*{Acknowledgements}

>>>>

Our basic approach is to extend predicates with two time parameters,
one corresponding with the starting time of the action and one with
its stopping time. For example, if John walked between $t_0$ and
$t_1$, we would have: $walk(john,t0,t1)$. From now on we will call an
interval of time points $[t_0,t_1]$ a timespan. Every timespan
$[t_0,t_1]$ has the property $t_0 ≤ t_1$ (it starts no later than it
stops).

\paragraph{temporal context}

We adjust the montegovian semantics so that every category takes a
timespan as an additional parameter. This timespan serves as a time
reference for semantics, and we call it a \emph{temporal context}. This current interval propagates through the compositional
interpretation down to atoms. By default, every interpretation passes
the temporal context down to its components without changing
it. However some key elements will act on it on nontrivial ways, which
we proceed to detail below.

\paragraph{tenses}

The prime non-trivial manipulation of timespans are tenses. In our
syntax, inherited from GF, tenses are represented syntactically as an
attribute of the clause. TODO{show example}

In our representation we deal only with present and simple past
tenses.  Indeed we find that FraCas does not exercise additional
specific tenses (if a more complicated tense is used, the additional
information is also carried by adverbs or adverbial phrases, in a more
specific way.)

TODO: optional time context.

The procedure to handle tense annotations is as follows:
\begin{itemize}
\item If the temporal context was modified by an adverb (or an
  averbial clause, such as ``before James swam''), then the tense does
  not create a new interval, but it may constrain it. Typically, a
  past tense adds the constraint that the temporal context ends before
  the constant $now$, which is a constant representing the current
  time.
\item Otherwise, if the tense is the present, then the temporal
  context becomes the simple $(now,now)$ interval.
\item
  If the tense is the past, then we locally quantify over a time
  interval $[t_0,t_1]$, such that $t_1 < now$.
\end{itemize}

\paragraph{temporal adverbs}

The other single most important source of interesting timespans are
adverbs. Most of the temporal adverbs fall in either of the following
categories:

\begin{enumerate}
\item[exact interval adverbs] For such adverbs, an exact interval is
  provided. Typical such adverbs specify in fact a single point in
  time (so the start and the end of the interval coincide).

  ⟦``at 5 pm, s''⟧(...) = ⟦s⟧(5pm,5pm)

\item[existential interval] The majority of temporal adverbs
  existentially quantify over a timespan. Examples include ``since
  1991'', ``in 1996'', ``for two years'', etc. The common theme is to
  introduce the interval and then restrict its bounds or its duration
  in someway. Sometimes the restriction is an equality (``for exactly
  two hours'').

  ⟦``in 1992, s''⟧(...) = ∃t1,t2.  [t1,t2] ⊆ 1992, ⟦s⟧(t1,t2)

\item[universal interval] A few adverbs introduce intervals via a
  universal quantification (sometimes with a constraint). Examples
  include: ``always'', ``never''.

  If there is no explicit time context, then ``always'' has no
  constraint on the interval, otherwise the quantified interval must
  be included in it:
  ⟦``always s''⟧(t0,t1) = ∀t0',t1'. [t0',t1'] ⊆ [t0,t1], ⟦s⟧(t0',t1')

  % A particularly interesting case found in FraCas is ``every month'':

  % ⟦``every month s''⟧(t0,t1) = ∀t. [t,t+1month] ⊆ [t0,t1], ⟦s⟧(t,t+1month)
\end{enumerate}

@@

A common theme in the testsuite is to refer to previous occurences of
an event:

(262)
P1	Smith left after Jones left.
P2	Jones left after Anderson left.
Q 	Did Smith leave after Anderson left?
H 	Smith left after Anderson left.

To be able to conclude as the testsuite expects, we have to make sure
that the two occurences of ``Jones left'', in P1 and P3 refer to the
same time intervals.

@@

Unicity of action appears to be a non-logical principle. Indeed, it is
quite possible that ``Jones left'' several times. However, it seems
that this principle is never contradicted by the testsuite. As such,
even though unicity of action is only a pragmatic rule, it can be
taken as a valid \emph{by default}: it is only when we have a
sufficiently constrained situation that one should reject it. Consider
the following discourse:

(1) Smith left at 1pm.
(2) Smith went to its appointment with the lawyer.
(3) Smith came back at 2pm
(4) Smith left at 4pm.

One would normally not say that there is contradiction. However if (1)
and (4) are presented in isolation, a contradiction can be flagged.

\paragraph{(un)repeatable achievements}
The principle of using unicity of action interacts well with the usual interpretation of quantifiers (and anaphora).
Indeed, using it, we can refute (279):

279:
P1	Smith wrote a novel in 1991.
Q 	Did Smith write it in 1992?

Indeed, following our account, the above inference problem is be
interperted as ∃x. novel(x) ∀t<NOW.  ∧ write(smith,x,t,t) [the scope
for the existential is extended according to our previous analysis of
anaphora]

@@

@@

\section{Results and evaluation}
\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}[hbt]
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#FraCaS
                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
Quantifiers  & 75         & .91        & .96         & .96    & .77  & .53  & .93  \\
             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
Plurals      & 33         & .64        & .82         & .76    & .67  & .52  & .73 \\
             &            &            &             &     &   &   & \ncases{24} \\
Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
Comparatives & 31         & .64        & .87         & .56    & .48  & .45  &  -       \\
Temporal     & 75         & .71        &  -          &   -    &   -  &  -   &  -       \\
Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
             &            &            &             &        &      &      & \ncases {9}  \\ \hline
Total        & 337        & .78        & .89         & .83    & .69  & .50  & .85  \\
             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Accuracy of our system compared to others.
    ``Ours" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FraCoq''
    refers to the work of \citet{bernardy:2017}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilizes the first-order automated theorem prover
    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
    to the system presented by \citet{Abzianidze:2015}. A dash
    indicates that no attempt was made for the section. }
  \label{tab:results}
\end{table}
\paragraph{interaction with negation and group readings}
\section{Conclusions and Future Work}
Stergios
\section*{Acknowledgements}
